---
title: "Hierarchical Data"
author: "Michael Helm"
date: "`r Sys.Date()`"
output: html_document
---

# Hierachical data

## Lists

Lists can hold different types of elemtents
```{r}
x1 <- list(1:4, "a", TRUE)
x1
```


We can often name the components or children of a list by naming the column of a tibble 

```{r}
x2 <- list(a = 1:2, b = 1:3, c = 1:4)
x2
```

The function strucure can be shown with `str()`

```{r}
str(x1)

str(x2)
```
## Hierarchy 

Lists can also include lists within them

```{r}
x3 <- list(list(1, 2), list(3, 4))
str(x3)
```

`C()` will create a flat vector. Function `str()` gets more useful as the the list gets more complex

```{r}
x5 <- list(1, list(2, list(3, list(4, list(5)))))
str(x5)
```
## List-Columns 

Lists can also live inside a tibble where we call them list-columns.


```{r}
library(tibble)
library(tidyr)
library(dplyr)

df <- tibble(x = 1:2, 
  y = c("a", "b"),
  z = list(list(1, 2), list(3, 4, 5))
)
df
```

## Unnesting


With a named list like exactly below, they would unnest into different columns. Below x would contain elements 1, 2, 3 and y would contain a list of 2 numbers that are named.

```{r}
df1 <- tribble(
  ~x, ~y,
  1, list(a = 11, b = 12),
  2, list(a = 21, b = 22),
  3, list(a = 31, b = 32),
)

df1 |> 
  unnest_wider(y)

df1 |> 
  unnest_wider(y, names_sep = "_")
```

Now if the children are unamed then they would unest into rows 

```{r}
df2 <- tribble(
  ~x, ~y,
  1, list(11, 12, 13),
  2, list(21),
  3, list(31, 32),
)

df2 |> 
  unnest_longer(y)
```

let's take a look of what happens when there is an empty element
```{r}
df6 <- tribble(
  ~x, ~y,
  "a", list(1, 2),
  "b", list(3),
  "c", list()
)
df6 |> unnest_longer(y)
```

There is also the fucntion of `unest_auto()`that chooses between longer and wider

## Case Studies 

gh_repos is a list, but our tools work with list-columns, so we’ll begin by putting it into a tibble.

```{r}
library(repurrrsive)

repos <- tibble(json = gh_repos)
repos
```

As you see the `gh_repos` is a list of a list. There are 6 list with 26 or 30 elements. We made it into a tibble. If we now use `unest_longer()` we put each child in it's own row. If we use both longer and wider than then we can put each element into its own column

```{r}
repos |> 
  unnest_longer(json)
```

Since each element is named list we can use `unnest_wider()` to put wach element into its column

```{r}
repos |> 
  unnest_longer(json) |> 
  unnest_wider(json) 
```

IF we want to find the names of each column
```{r}
repos |> 
  unnest_longer(json) |> 
  unnest_wider(json) |> 
  names() |> 
  head(10)
```

```{r}
repos |> 
  unnest_longer(json) |> 
  unnest_wider(json) |> 
  select(id, full_name, owner, description) |> 
  unnest_wider(owner, names_sep = "_") ##if we stop here there is an error since owner is 
## a list that contains a cloumn id so we add names_sep to account for that
```

## Relational data

Below is data about character in the Game of Thrones
```{r}
chars <- tibble(json = got_chars)

chars %>%
  unnest_wider(json) %>%
  select(id, name, gender, culture, born, died, alive)
```

Exploring the titles column
```{r}
chars |> 
  unnest_wider(json) |> 
  select(id, titles) |> 
  unnest_longer(titles)
```


Using hoist will allow a



## JSON

JSON is short for javascript object notation and is the common way web API return data. Was made to be easily read by computers but not humans.

There our six key data types. Four of them are scalars.
- sting
- number
- boolean 
- null

There is a an
- arrary, an unamed list in []
- object, is like a named list in {}, {'c' = 2}

Json does not have a native way to represent dates or date-times. Will need to use
- readr::parse_date()
- readr::parse_datetime()
- readr::parse_double()

### jsonlite

When in R we will use the `jsonlite` package. We will use functions. 

- `read_json()` and `parse_json()`
- in real lide youll use read_json to read a JSON file from disk
- The repurrsive package `gh_user` used earlier can be read in by `read_json()`

```{r}
library(jsonlite)
library(purrr)

gh_users_json()

gh_user2 <- read_json(gh_users_json())
```

```{r}

str(parse_json('1'))

str(parse_json('[1, 2, 3]'))

str(parse_json('{"x": [1, 2, 3]}'))

```



In most cases JSON files contain a sinfle top level array, because they're designed to provide data about multiple things

```{r}
json <- '[
  {"name": "John", "age": 34},
  {"name": "Susan", "age": 27}
]'
df <- tibble(json = parse_json(json))
df


df |> 
  unnest_wider(json)

```

If JSON file consists of a sinfle top-level JSON object, representing one "thing. We start the rectangling process by wrapping it in a list, before you put it in a tibble.



```{r}
json <- '{
  "status": "OK", 
  "results": [
    {"name": "John", "age": 34},
    {"name": "Susan", "age": 27}
 ]
}
'
df <- tibble(json = list(parse_json(json)))
df


df |> 
  unnest_wider(json) |> 
  unnest_longer(results) |> 
  unnest_wider(results)

```

or

```{r}
df <- tibble(results = parse_json(json)$results)
df |> 
  unnest_wider(results)
```

## Questions

What is a list vs tibble vs dataframe?

How can we tell a single top-level array representing multiple things vs. a single top-level array representing a single thing?

Rectangling Process? 

Tibble vs tribble?

